---
title: "linear model from Web"
author: "Michael Maida"
date: "Fall 2021"
output:
   html_document:
         toc: true
         toc_depth: 5
        
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read about the data from the website where it is staged.  Then form three regression models; Model1 is SIMS~ARM, Model2 is SIMS~GRIP, and Model3 is SIMS~ARM+GRIP. For each model find a 95% prediction interval of SIMS given a value of 94  for GRIP and 88 for ARM. Compare Model1 with Model3 using anova. Write it up in a markdown document, push the project up to your github account and submit it back to canvas as link. 

 


```{r include=FALSE}
require(tidyverse)
require(tigerstats)
require(rgl)
require(knitr)

```

```{r}
data <- read.csv(file="https://onlinestatbook.com/case_studies_rvls/physical_strength/data.txt",sep="",header=TRUE)  

```

## Model 1 SIM~ARMS

### scatterplot

```{r}
plot(SIMS~ARM,data=data)
```

This SIMS vs ARM scatterplot indicates a correlation between the two variables. The graph shows an uphill trend; as ARM increases, SIMS increases. It is a moderately tight scatterplot, with a relatively consistent bond width of around 4. Since it represents a normal distribution, and normal distributions have 4 standard deviations from the top of the bond width to the bottom of the bond width, one standard deviation could be measured at around 1 from top to bottom. 

### Numerical results

```{r}
cor(SIMS~ARM,data=data)
```
The correlation is 0.6860073, which is fairly well coordinated.

### Inferential  (Build model.1)

```{r}
model.1 <- lm(SIMS~ARM,data=data)
summary.lm(model.1)
```
The equation of the line is:   
SIMS = m * ARMS + b   
SIMS = 0.054563 * ARMS - 4.095160   

Our adjusted R-squared is 0.467, which indicates that ARM explains under half of the variation of the SIMS data. In other words, it reduces the error the the mean model by 46.7%. The residual standard error was calculated at 1.226, so the standard deviation is 1.226. This means that the bond width is 4.904, which is decently narrow. This indicates that ARMS can likely be used to predict SIMS.

#### Predict at target point

```{r}
SIMS = 0.054563 * 88 - 4.095160  
mydata = data.frame(GRIP = 94, ARM=88)
predict(model.1, mydata, interval="prediction")
```

The 95% confidence interval goes from -1.726 to 3.139--a distance of 4.865. This is 2 SD's up and 2 SD's down. Since the residual standard error is 1.226, which is approximately one fourth of the 95% confidence interval. 

#### scatterplot with model fit

```{r}
basicNN <- ggplot(data, aes(y=SIMS,x=ARM))
```

```{r}
basicNN + geom_point() + geom_smooth(method=lm)
``` 

Here is our linear fit. At any point above the line is a normal curve with a standard deviation of 1.226.

## Model 2 SIM~GRIP

### Now add in scatterplot


### Numerical results 


### Inferential  (Build model.2)
  

#### predict model.2 at target point

#### now add the model fit to our plot for model.2
  


## Model 3 SIM~ARM+GRIP

### Numerical results (cor)

`
  
### Inferential  (Build 2-dimentional model.3)

#### predict model.3 at target point
  

## Comparing nested models ANOVA Test

### Model.1 vs Model.3
### Model.2 vs Model.3

## Informally compare Model.1 with model.2

