---
title: "linear model from Web"
author: "Michael Maida"
date: "Fall 2021"
output:
   html_document:
         toc: true
         toc_depth: 5
        
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read about the data from the website where it is staged.  Then form three regression models; Model1 is SIMS~ARM, Model2 is SIMS~GRIP, and Model3 is SIMS~ARM+GRIP. For each model find a 95% prediction interval of SIMS given a value of 94  for GRIP and 88 for ARM. Compare Model1 with Model3 using anova. Write it up in a markdown document, push the project up to your github account and submit it back to canvas as link. 

 


```{r include=FALSE}
require(tidyverse)
require(tigerstats)
require(rgl)
require(knitr)

```

```{r}
data <- read.csv(file="https://onlinestatbook.com/case_studies_rvls/physical_strength/data.txt",sep="",header=TRUE)  

```

## Model 1 SIM~ARMS

### scatterplot

```{r}
plot(SIMS~ARM,data=data)
```

This SIMS vs ARM scatterplot indicates a correlation between the two variables. The graph shows an uphill trend; as ARM increases, SIMS increases. It is a moderately tight scatterplot, with a relatively consistent bond width of around 4. Since it represents a normal distribution, and normal distributions have 4 standard deviations from the top of the bond width to the bottom of the bond width, one standard deviation could be measured at around 1 from top to bottom. 

### Numerical results

```{r}
cor(SIMS~ARM,data=data)
```
The correlation is 0.6860073, which is fairly well coordinated.

### Inferential  (Build model.1)

```{r}
model.1 <- lm(SIMS~ARM,data=data)
summary.lm(model.1)
```

The equation of the line is:   
SIMS = m * ARMS + b   
SIMS = 0.054563 * ARMS - 4.095160   

Our adjusted R-squared is 0.467, which indicates that ARM explains under half of the variation of the SIMS data. In other words, it reduces the error of the mean model by 46.7%. The residual standard error was calculated at 1.226, so the standard deviation is 1.226. This means that the bond width is 4.904, which is decently narrow. This indicates that ARMS can likely be used to predict SIMS.

#### Predict at target point

```{r}
SIMS = 0.054563 * 88 - 4.095160  
mydata = data.frame(GRIP = 94, ARM=88)
predict(model.1, mydata, interval="prediction")
```

The 95% confidence interval goes from -1.726 to 3.139--a distance of 4.865. This is 2 SD's up and 2 SD's down. Since the residual standard error is 1.226, which is approximately one fourth of the 95% confidence interval, the . 

#### scatterplot with model fit

```{r}
basicNN <- ggplot(data, aes(y=SIMS,x=ARM))
```

```{r}
basicNN + geom_point() + geom_smooth(method=lm)
``` 

Here is our linear fit. At any point above the line is a normal curve with a standard deviation of 1.226.

## Model 2 SIM~GRIP

### Now add in scatterplot

```{r}
plot(SIMS~GRIP,data=data)
```

Just as in the former case, there seems to be an uphill trend to the right, indicative of a positive linear correlation. 

### Numerical results 

```{r}
cor(SIMS~GRIP,data=data)
```

The correlation in the prior example was 0.6860073, while the correlation with SIMS~GRIP is only 0.6398458. Still, this is a fairly large correlation.

### Inferential  (Build model.2)

```{r}
model.2 <- lm(SIMS~GRIP,data=data)
summary.lm(model.2)
```

The equation of the line is:   
SIMS = m * ARMS + b   
SIMS = 0.045463 * ARMS - 4.809675

Our adjusted R-squared is 0.4053, which indicates that GRIP explains less than half of the variation of the SIMS data. It reduces the error of the mean model by 40.53%. The residual standard error was calculated at 1.295, which is equal to the standard deviation. This means the bond width is 1.295 * 4, or 5.18. Because the bond width is so narrow, it can be used to predict SIMS. But because it is greater than the previous example with ARM, it is less accurate in its predictions.

#### predict model.2 at target point

```{r}
predict(model.2, mydata, interval="prediction")
```

Here, the prediction of GRIP is 94 when the prediction of ARM is -0.536. The difference between the lower and upper limits is 5.143613, around 5.14.

#### now add the model fit to our plot for model.2
  
```{r}
basicNT <- ggplot(data, aes(y=SIMS,x=GRIP))
basicNT + geom_point() + geom_smooth(method=lm)
``` 

Here is our linear fit for SIMS~GRIP. At any point above the line is a normal curve with a standard deviation of 1.295. 

## Model 3 SIM~ARM+GRIP

### Numerical results (cor)

```{r}
cor(SIMS~ARM+GRIP,data=data)
```

The correlation between SIMS~ARM was 0.6860073. The correlation between SIMS~GRIP was  0.6398458. If these two correlations were considered fairly large, then the correlation between SIMS and ARM+GRIP is even greater, standing at 0.7326545. This is a notably large correlation.

### Inferential  (Build 2-dimentional model.3)

```{r}
model.3 <- lm(SIMS~GRIP + ARM,data=data)
summary.lm(model.3)
``` 
SIMS = (0.024470 * GRIP) + (0.037311 * ARM) - 5.433871

#### predict model.3 at target point

```{r}
predict(model.3, mydata, interval="prediction")
```

When ARM is 88 and GRIP is 94, the prediction of SIMS is 0.149. The 95% confidence interval goes from -2.132373 to 2.431668--a distance of 4.564041. This is 2 standard deviations up and 2 standard deviations down. Since the residual standard error is 1.144, which is approximately one fourth of the 95% confidence interval.

## Comparing nested models ANOVA Test

### Model.1 vs Model.3

```{r}
anova(model.1,model.3)
```

Model 1, which only depicts SIMS as explained by ARM, has an amount of error equal to 217.88. When SIMS was explained by GRIP as well as ARM, the amount of error decreased to 188.43. This shows that there is indeed a significant difference between the two models. Thus, we accept the null hypothesis since there is a significant statistical difference. 

### Model.2 vs Model.3

```{r}
anova(model.2,model.3)
```

Model 1 shows the amount of error in SIMS explained by GRIP, which was equal to 243.07. In contrast, SIMS explained by GRIP + ARM has an amount of error of 188.43, which is significantly lower. Thus, we accept the null hypothesis: there is a clear statistical difference between the two models.

## Informally compare Model.1 with model.2

```{r}
anova(model.1,model.2)
```

While no P-values are available since they are nested models, you can still informally compare them. The amount of error between Model 1 and 2 is 25.19. This indicates that ARM better explains SIMS than GRIP, since it has a noticeably lesser amount of error. 